
\newpage

\chapter*{Acknowledgement}
\phantomsection
\addcontentsline{toc}{chapter}{Acknowledgement}

\begin{center} 
    \begin{minipage}{0.75\textwidth}
        \begin{quote}\itshape
            “The best way to predict the future is to create it.”\par
            \hfill— Alan Kay
          \end{quote}
       \begin{quote}\itshape
            “If we knew what it was we were doing, ...”\par
             \hfill— Albert Einstein
       \end{quote}
          \begin{quote}\itshape 
            "Will we go down? As far as we know,
            How do we know?"\par
            \hfill— Anna Von Hausswolff
          \end{quote}
    \end{minipage}
\end{center}
\vspace{0.5cm}

The following study is a bachelor thesis in Software Engineering at Jönköping University. As not uncommon for the female nature, behavioral 
patterns interests us, beyond doubt in combination with technology potentially holdning capacity to change not only society, but how we think. 
For the latter, it should be acknowledged, I surely hope not. 

\medskip
Thanks to Knowit, Jönköping, for supporting our project even if interdisciplinary to the general software development area. 
Thanks to our beloved computers for bringing some light into the dark night, consuming our eyes and sometimes let us forget when our brain fight. 
Thanks to all participants in this study, sharing your thoughts and a fragment of your minds, after all, it is what us all binds. 
Thanks to my thesis partner, and dear friend, Janna, you are worth more than a dime. 
Glad you read my rhyme, purhaps some joy is found when reading what is following, at least part-time. 

\medskip
Jönköping, May 2025

\medskip
Sara Ljung

\vspace{0.3cm}


